---
title: "Assignment-3"
author: "Chirag Vithalani"
date: "September 10, 2016"
output: html_document
---

<hr />



#### Page 113: #2
The following table gives the elongation e in inches per inch (in./in.) for a given stress
S on a steel wire measured in pounds per square inch $(lb/in^2)$. Test the model e=c<sub>1</sub>S
by plotting the data. Estimate c<sub>1</sub> graphically.

```{r warning=FALSE}
library(reshape2)
library(ggplot2)

S<- c(5,10,20,30,40,50,60,70,80,90,100)
e<- c(0,19,57,94,134,173,216,256,297,343,390)

dataframe<- data.frame(S,e)

plot1<- ggplot(data = dataframe, aes(x = dataframe$S, y = dataframe$e)) + geom_point()

plot1 <- plot1 + labs(title="Data", x="S", y="e")
plot1



```


The value of c<sub>1</sub>=(390-343)/(100-90)= 4.7

so equation can be written as e=4.7S

#### Page 121: #2.a
For each of the following data sets, formulate the mathematical model that minimizes
the largest deviation between the data and the line y=ax+b. If a computer is available,
solve for the estimates of a and b.

```{r warning=FALSE}

x<- c(1.0, 2.3, 3.7, 4.7, 6.1, 7.0)
y<- c(3.6, 3.0, 3.2, 5.1, 5.3, 6.8)

dataframe<- data.frame(x,y)

plot2<- ggplot(data = dataframe, aes(x = x, y = y)) + geom_line()
plot2 <- plot2 + labs(title="Data Points", x="x", y="y")
plot2

```

If we choose (x, y) = (6.1, 5,3) and (x, y) = (7.0, 6.8) ponts to find the slope a and the y-intercept b:


```{r}
x<- c(1.0, 2.3, 3.7, 4.7, 6.1, 7.0)
y<- c(3.6, 3.0, 3.2, 5.1, 5.3, 6.8)
dataframe<- data.frame(x,y)


dataframe$z = c(NA,tail(dataframe$y,-1)-head(dataframe$y,-1))
dataframe$z2= c(NA,tail(dataframe$x,-1)-head(dataframe$x,-1))

dataframe$slope<- dataframe$z / dataframe$z2

slope<- mean(dataframe$slope, na.rm= TRUE)
slope

xmean<- mean(dataframe$x)
ymean<- mean(dataframe$y)

b<- ymean - xmean*slope
b

```
so our equation would be y=0.67x+1.69

```{r}
dataframe$fx<- (.67*dataframe$x) + 1.69

plot3<- ggplot(data = dataframe, aes(x = x, y = fx)) + geom_line()

plot3 <- plot3 + labs(title="y=ax+b", x="x", y="y")
plot3
```


#### Page 127, #10 

```{r}

planets <- data.frame(body = c("Mercury", "Venus", "Earth", "Mars", "Jupiter", 
                               "Saturn", "Uranus", "Neptune"), 
                      period = c(7.6e6, 1.94e7, 3.16e7, 5.94e7, 3.74e8, 9.35e8, 2.64e9, 5.22e9),
                      distance = c(5.79e10, 1.08e11, 1.5e11, 2.28e11, 7.79e11, 1.43e12, 
                                   2.87e12, 4.5e12))



```

Fit the model $y = ax^3/2$

In this case, our formula is:

$$ a = \frac{\Sigma x_i^n y_i}{\Sigma x_i^{2n}} $$



```{r}
a <- sum(planets$period^(3/2)*planets$distance)/sum(planets$period^(6/2))

a

```

So equation  $distance = 0.01320756*x^{\frac{3}{2}}$

#### Page 157, # 4 

In the following data, X represents the diameter of a Ponderosa Pine measured at breast height, and 
Y is a measure of volume - number of bard feet divided by 10. Make a scatterplot of the data.Discuss the appropriateness of using a 13th degree polynomial that passes through the data points as an empirical model.If you have a computer available, fit a polynomial to the data and graph the results.


We have 14 datapoints, a 13th degree polynomial exists that would pass through every data 
point. However, this would be an example of overfitting: it would fit the data perfectly, but 
woudln't be applicable to new observations.


```{r}

ponderosa <- data.frame(x = c(17, 19, 20, 22, 23, 25, 31, 32, 33, 36, 37, 38, 39, 41),
                        y = c(19, 25, 32, 51, 57, 71, 141, 123, 187, 192, 205, 252, 248, 294))

library(ggplot2)

ggplot(ponderosa, aes(x=x, y=y)) + geom_point()

```


13th degree polynomial that passes through the data points as an empirical model means 14 data points require that the satisfy the list of linear algebraic equations.

We can use poly() function and the fitting linear model function lm() to solve.

```{r}

summary(lm(y ~ poly(x, 13, raw=TRUE), data=ponderosa))

coefficients <- as.vector(lm(y ~ poly(x, 13, raw=TRUE), data=ponderosa)$coefficients)

coefficients # some values have value of "NA""

ggplot(ponderosa) + geom_point(aes(x=x, y=y)) + geom_line(aes(x=x, y=
  coefficients[1] + coefficients[2]*x + coefficients[3]*x^2 +
    coefficients[4]*x^3 + coefficients[5]*x^4 + coefficients[6]*x^5 + 
    coefficients[7]*x^6 + coefficients[8]*x^7 + coefficients[9]*x^8 +
    coefficients[11]*x^10 +
    coefficients[13]*x^12))

```

In above plot we can see that this is not exactly lagrangian polynomial but this has ability to capture the trends of data collected. But small change in error will cause high order of movement back and forth change in the data so this model is not too useful in modelling. 
