---
title: "Assignment-1"
author: "Chirag Vithalani"
date: "September 10, 2016"
output: html_document
---

<hr />
####Page 8: #10

Your grandparents have an annuity. The value of the annuity increases each month by
an automatic deposit of 1% interest on the previous month's balance.Your grandparents
withdraw $1000 at the beginning of each month for living expenses. Currently, they
have $50,000 in the annuity. Model the annuity with a dynamical system. Will the
annuity run out of money? When? Hint: What value will 
<math xmlns="http://www.w3.org/1998/Math/MathML">
<msub><mi>a</mi><mi>n</mi></msub>
</math> have when the annuity is
depleted?

As per on page 6

<math xmlns="http://www.w3.org/1998/Math/MathML"><mi>c</mi><mi>h</mi><mi>a</mi><mi>n</mi><mi>g</mi><mi>e</mi><mo>=</mo><mo>&#x2206;</mo><mo>&#xA0;</mo><msub><mi>a</mi><mi>n</mi></msub><mo>=</mo><msub><mi>a</mi><mrow><mi>n</mi><mo>+</mo><mn>1</mn></mrow></msub><mo>-</mo><mo>&#xA0;</mo><msub><mi>a</mi><mi>n</mi></msub></math>

<math xmlns="http://www.w3.org/1998/Math/MathML">
<msub><mi>a</mi><mi>0</mi></msub>
</math>=50,000

Initial value is $50,000
Now after first month it will be 
$50,000 (initial value) + $500 ( 1% interest) -$1000 ( withdrawal)=$49500=<math xmlns="http://www.w3.org/1998/Math/MathML">
<msub><mi>a</mi><mi>1</mi></msub>
</math>




<math xmlns="http://www.w3.org/1998/Math/MathML"><mi>c</mi><mi>h</mi><mi>a</mi><mi>n</mi><mi>g</mi><mi>e</mi><mo>=</mo><mo>&#x2206;</mo><mo>&#xA0;</mo><msub><mi>a</mi><mi>n</mi></msub><mo>=</mo><msub><mi>a</mi><mrow><mi>n</mi><mo>+</mo><mn>1</mn></mrow></msub><mo>-</mo><msub><mi>a</mi><mi>n</mi></msub><mo>=</mo><mn>0</mn><mo>.</mo><mn>01</mn><msub><mi>a</mi><mi>n</mi></msub><mo>-</mo><mn>1000</mn><mspace linebreak="newline"/><mi>s</mi><mi>o</mi><mspace linebreak="newline"/><msub><mi>a</mi><mrow><mi>n</mi><mo>+</mo><mn>1</mn></mrow></msub><mo>-</mo><msub><mi>a</mi><mi>n</mi></msub><mo>=</mo><mn>0</mn><mo>.</mo><mn>01</mn><msub><mi>a</mi><mi>n</mi></msub><mo>-</mo><mn>1000</mn><mspace linebreak="newline"/><mi>f</mi><mi>u</mi><mi>n</mi><mi>c</mi><mi>t</mi><mi>i</mi><mi>o</mi><mi>n</mi><mo>&#xA0;</mo><mi>w</mi><mi>o</mi><mi>u</mi><mi>l</mi><mi>d</mi><mo>&#xA0;</mo><mi>b</mi><mi>e</mi><mspace linebreak="newline"/><msub><mi>a</mi><mrow><mi>n</mi><mo>+</mo><mn>1</mn></mrow></msub><mo>=</mo><msub><mi>a</mi><mi>n</mi></msub><mo>+</mo><mn>0</mn><mo>.</mo><mn>01</mn><msub><mi>a</mi><mi>n</mi></msub><mo>-</mo><mn>1000</mn><mo>=</mo><msub><mi>a</mi><mi>n</mi></msub><mo>(</mo><mn>1</mn><mo>+</mo><mn>0</mn><mo>.</mo><mn>01</mn><mo>)</mo><mo>-</mo><mn>1000</mn><mo>=</mo><mn>1</mn><mo>.</mo><mn>01</mn><msub><mi>a</mi><mi>n</mi></msub><mo>-</mo><mn>1000</mn></math>

```{r}

amountLeft=50000 #Initial amount
month=0 #initially 0th month

while(amountLeft >0)
{
  month=month+1
  
  amountLeft=1.01 * amountLeft-1000
  
  cat(sprintf("Amount left: %f after %d month \n", amountLeft, month))
  
    
}

```
####so annuity will run out after 69 months


####Page 17: #9
The data in the accompanying table show the speed n (in increments of 5 mph) of an
automobile and the associated distance an in feet required to stop it once the brakes
are applied. For instance, n=6 (representing 6 x 5 = 30 mph) requires a stopping
distance of <math xmlns="http://www.w3.org/1998/Math/MathML"><msub><mi>a</mi><mn>6</mn></msub></math>=47ft.

a. Calculate and plot the change $\Delta a_n$ versus $n$. Does the graph reasonably approximate a 
linear relationship?

<math xmlns="http://www.w3.org/1998/Math/MathML" >
  <mi mathvariant="normal">&#x0394;<!-- ?? --></mi>
  <msub>
    <mi>a</mi>
    <mi>n</mi>
  </msub>
  <mo>=</mo>
  <msub>
    <mi>a</mi>
    <mrow >
      <mi>n</mi>
      <mo>+</mo>
      <mn>1</mn>
    </mrow>
  </msub>
  <mo>&#x2212;<!-- ??? --></mo>
  <msub>
    <mi>a</mi>
    <mi>n</mi>
  </msub>
</math>

<math xmlns="http://www.w3.org/1998/Math/MathML" >
  <mi mathvariant="normal">&#x0394;<!-- ?? --></mi>
  <msub>
    <mi>a</mi>
    <mn>0</mn>
  </msub>
  <mo>=</mo>
  <msub>
    <mi>a</mi>
    <mn>1</mn>
  </msub>
  <mo>&#x2212;<!-- ??? --></mo>
  <msub>
    <mi>a</mi>
    <mn>0</mn>
  </msub>
  <mo>=</mo>
  <mn>6</mn>
  <mo>&#x2212;<!-- ??? --></mo>
  <mn>3</mn>
  <mo>=</mo>
  <mn>3</mn>
</math>

```{r warning=FALSE}

brakesDataFrame <- data.frame(n=1:16, an=c(3,6,11,21,32,47,65,87,112,140,171,204,241,282,325,376))

brakesDataFrame$deltaAn = c(NA,tail(brakesDataFrame$an,-1)-head(brakesDataFrame$an,-1))
brakesDataFrame

library(ggplot2)

ggplot(brakesDataFrame, aes(x=n, y=deltaAn)) +geom_line()

```

This is almost linear relationship.

b. Based on your conclusions in part (a), find a difference equation model for the stopping distance
data. Test your model by plotting the errors in the predicted values against $n$. Discuss the 
appropriateness of the model.

we can assume that if n=0, stopping distance will be 0, that is line passes through origin.

To approximate the slope of this line calculating $\frac{\Delta a_n}{n}$



```{r}

brakesDataFrame$slope <- brakesDataFrame$deltaAn / brakesDataFrame$n

slope <- mean(brakesDataFrame$slope, na.rm=TRUE)
slope

```


Plotting errors in the predicted values vs n

```{r}

brakesDataFrame$estimate <- slope*brakesDataFrame$n # as per page 3

brakesDataFrame$error <- brakesDataFrame$estimate - brakesDataFrame$deltaAn

ggplot(brakesDataFrame, aes(x=n, y=error)) + geom_point()


```

Looks like there is pattern to the errors. This model may not be appropriate.


### Page 34, Question 13 ###

Consider the spreading of a rumor through a company of 1000 employees, all working in the same 
building. We assume that the spreading of a rumor is similar to the spreading of a contagious 
disease (example 3, section 1.2) in that the number of people hearing the rumor each day is 
proportional to the product of the number who have heard the rumor previously and the number who
have not heard the rumor. This is given by:

$$ r_{n+1} = r_n + kr_n (1000 - n) $$

Where $k$ is a parameter that depends on how fast the rumor spreads and $n$ is the number of days.
Assume $k = 0.001$ and further assume that four people initially have heard the rumor. How soon will
all 1000 employees have heard the rumor?

####Answer
$n$ is the number of days
assume that four people initially have heard the rumor
```{r}


rumorDataFrame <- data.frame(n=0, r=4)

k <- 0.001

for(i in 1:13)
{
  # as per given equation
  tempvar<-c(i, rumorDataFrame$r[i] + k*rumorDataFrame$r[i]*(1000 - rumorDataFrame$r[i]))
  rumorDataFrame <- rbind(rumorDataFrame,tempvar )
}

rumorDataFrame

```

We can say,after 12 days, the rumor will have spread throughout the office.

### Page 55, Question 6 ###

An economist is interested in the variation of the price of a single product. It is observed that a 
high price for the product in the market attracts more suppliers. However, increasing the quantity
of the product supplied tends to drive the price down. Over time, there is an interaction between 
price and supply. The economist has proposed the following model,p where $P_n$ represents the price 
of the product at year $n$, and $Q_n$ represents the quantity. Find the equilibrium values for this
system.

$$ P_{n+1} = P_n - 0.1(Q_n - 500) $$
$$ Q_{n+1} = Q_n + 0.2(P_n - 100) $$

a. Does the model make sense intuitively? What is the significance of the constants
100 and 500? Explain the significance of the signs of the constants -0.1 and 0.2.???0:1 and 0.2.

Yes it looks OK intuitively as it simpliefies things :)


100 and 500 looks equilibrium values.
For that
<math xmlns="http://www.w3.org/1998/Math/MathML"><mi>P</mi><mi>n</mi><mo>=</mo><mi>P</mi><mi>n</mi><mo>+</mo><mn>1</mn><mo>&#xA0;</mo><mo>&#xA0;</mo><mi>a</mi><mi>n</mi><mi>d</mi><mo>&#xA0;</mo><mi>Q</mi><mi>n</mi><mo>=</mo><mi>Q</mi><mi>n</mi><mo>+</mo><mn>1</mn></math>

If we put values

<math xmlns="http://www.w3.org/1998/Math/MathML"><mi>P</mi><mo>-</mo><mi>P</mi><mo>=</mo><mo>-</mo><mn>0</mn><mo>.</mo><mn>1</mn><mo>(</mo><mi>Q</mi><mo>-</mo><mn>500</mn><mo>)</mo><mspace linebreak="newline"/><mn>0</mn><mo>=</mo><mi>Q</mi><mo>-</mo><mn>500</mn><mspace linebreak="newline"/><mi>Q</mi><mo>=</mo><mn>500</mn></math>

sign looks the negative and positive price-quantity relationship.

b. Test the initial conditions in the following table and predict the long-term behavior:

     Case | Price | Quantity
----------|-------|---------
Case A    | 100   | 500
Case B    | 200   | 500
Case C    | 100   | 600
Case D    | 100   | 400


```{r warning=FALSE}

library(reshape2)

years <- 50

pinit <- 100
qinit <- 500

economistDataFrame <- data.frame(n = 0, p = pinit, q= qinit)

for(i in 1:years){
  economistDataFrame <- rbind(economistDataFrame,
                  c(i, tail(economistDataFrame$p,1) - 0.1*(tail(economistDataFrame$q,1) - 500),
                    tail(economistDataFrame$q,1) + 0.2*(tail(economistDataFrame$p,1) - 100)))
}

ggplot(melt(economistDataFrame, id="n"), aes(x=n, y=value, color=variable)) + geom_line()

#-------------------------------------------------------

pinit <- 200
qinit <- 500

economistDataFrame <- data.frame(n = 0, p = pinit, q= qinit)

for(i in 1:years){
  economistDataFrame <- rbind(economistDataFrame,
                  c(i, tail(economistDataFrame$p,1) - 0.1*(tail(economistDataFrame$q,1) - 500),
                    tail(economistDataFrame$q,1) + 0.2*(tail(economistDataFrame$p,1) - 100)))
}

ggplot(melt(economistDataFrame, id="n"), aes(x=n, y=value, color=variable)) + geom_line()

#-------------------------------------------------------
pinit <- 100
qinit <- 600

economistDataFrame <- data.frame(n = 0, p = pinit, q= qinit)

for(i in 1:years){
  economistDataFrame <- rbind(economistDataFrame,
                  c(i, tail(economistDataFrame$p,1) - 0.1*(tail(economistDataFrame$q,1) - 500),
                    tail(economistDataFrame$q,1) + 0.2*(tail(economistDataFrame$p,1) - 100)))
}

ggplot(melt(economistDataFrame, id="n"), aes(x=n, y=value, color=variable)) + geom_line()

#-------------------------------------------------------

pinit <- 100
qinit <- 400

economistDataFrame <- data.frame(n = 0, p = pinit, q= qinit)

for(i in 1:years){
  economistDataFrame <- rbind(economistDataFrame,
                  c(i, tail(economistDataFrame$p,1) - 0.1*(tail(economistDataFrame$q,1) - 500),
                    tail(economistDataFrame$q,1) + 0.2*(tail(economistDataFrame$p,1) - 100)))
}

ggplot(melt(economistDataFrame, id="n"), aes(x=n, y=value, color=variable)) + geom_line()

```

I do not think this is good equation to indicate relationship. If we change quality and price by half then graph doesn't keep relationship and goes mad.

```{r warning=FALSE}
pinit <- 50
qinit <- 200

economistDataFrame <- data.frame(n = 0, p = pinit, q= qinit)

for(i in 1:years){
  economistDataFrame <- rbind(economistDataFrame,
                  c(i, tail(economistDataFrame$p,1) - 0.1*(tail(economistDataFrame$q,1) - 500),
                    tail(economistDataFrame$q,1) + 0.2*(tail(economistDataFrame$p,1) - 100)))
}

ggplot(melt(economistDataFrame, id="n"), aes(x=n, y=value, color=variable)) + geom_line()

```
